---
title: "Run Many Simulations at Once"
author: "Matthew Parker"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

The purpose of this document is to explain how to automate creation of a list of parameter combinations, which can then be used to run simulations.

## Illustrative Example

To aid in understanding, we will consider the simple example of simulating a Binomial random variable with many different values for $N$ and $p$.

\[
  X \sim Bin(N,p)
\]

We will also specify $n$, the number of samples we want to generate with each parameter combination (in this case we will just use one value for $n$, but we could have many values just like for the other parameters).

## List Parameter Values

```{r parameterValues}
N_opts <- seq(10,50,10)
p_opts <- seq(0.1,0.9,0.2)
n_opts <- 100

param_list <- list(N=N_opts, p=p_opts, n=n_opts)
print(param_list)
```

## Enumerate Combinations

Create a data.frame where each row is a combination of parameters, and all parameter combinations are present in the data.frame.

```{r enumerateCombinations}
enum <- expand.grid(param_list, stringsAsFactors = FALSE)
head(enum)
nrow(enum) # print the number of rows in enum, need to know this for SLURM script
```

## Save For Later Use

By saving this to a file, we can access it when running a simulation study. We will save it to the file *df_enum_pops.RDS*.

```{r save_pops_rds, eval=FALSE}
saveRDS(object = enum, file = "./df_enum_pops.RDS")
```


## R Simulation Script

This is the script that will run the simulation. Save the script as *simulation.R*. Note that the counter variable is passed through the command line via the SLURM script (see the next section for details). The counter variable is used to select which row of the enumerated parameter combinations we want to run the simulation for.

Results will be appended to a csv file called *results.csv*

```{r simulationStep, eval=FALSE}
# store the command line arguments
args <- commandArgs(TRUE)
# create a counter for your index from the first command line argument
counter <- as.integer(args[1])
# load the data.frame we generated previously
enum <- readRDS("./df_enum_pops.RDS")
# MAX_COUNTER tells us when we are on the final simulation, and should 
# concatenate our csv files into one
MAX_COUNTER <- nrow(enum) 

# run simulation based on row "counter" of the enum data.frame
results <- rbinom(n    = enum[counter,]$n, 
                  size = enum[counter,]$N, 
                  prob = enum[counter,]$p)

# put results into a named vector (so you know where the data came from later!)
named_results <- c(n       = enum[counter,]$n, 
                   size    = enum[counter,]$N, 
                   prob    = enum[counter,]$p,
                   results = results)
results2 <- matrix(named_results, nrow=1)
colnames(results2) <- names(named_results)

# save the results to a csv file called results_i.csv, where i equals counter
# to protect against simultaneous file writing, we will concatenate the files at the end
write.table(x = results2, file = paste0("./results_",counter,".csv"), 
            append = F, sep = ",", row.names = F, col.names = T)

# wait until all csv files have been written, then concatenate
if(counter==MAX_COUNTER) {
  l        <- 0
  csvFiles <- NULL
  # wait until all files exist
  while(l < MAX_COUNTER) { 
    # read in csv files starting with "results_" and ending with ".csv"
    csvFiles = list.files(pattern="^results_.*\\.csv$")
    l <- length(csvFiles)
    # check if all files exist once every 10 seconds
    Sys.sleep(10) 
  }
  # combine all csv files into one
  onecsv = do.call(rbind, lapply(csvFiles, function(x) read.csv(x, stringsAsFactors = FALSE)))
  # write "results.csv"
  write.table(x = onecsv, file = paste0("./results.csv"), 
            append = F, sep = ",", row.names = F, col.names = T)
}
```

## SLURM SCRIPT

This is the script that will submit and run each simulation, save it in a file called *submit_script.sl*, you can run it using the command:

```{r ex1, eval=FALSE}
  sbatch submit_script.sl
```

We use the *--array=1-25* to run each of the simulations (note that 25 is the number of rows in our *df_enum_pops.RDS* file).

```{r slurmExample, eval=FALSE}
#!/bin/bash -l
#SBATCH -J run_name
#SBATCH --array=1-25
#SBATCH --time=0--01:00:00
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --cpus-per-task=1
#SBATCH --mem=1000M
#SBATCH --mail-user=mrparker909@gmail.com
#SBATCH --mail-type=ALL

module load r/3.3.3
cd $SLURM_SUBMIT_DIR
Rscript ./simulation.R $SLURM_ARRAY_TASK_ID >&
./log/log_${SLURM_JOB_ID}_${SLURM_ARRAY_TASK_ID}.txt
```






